\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Statement of Intent}
\author{Matthew Gardner}
\date{}

\begin{document}
\maketitle

I started out my post-secondary education double majoring at BYU in physics and
music.  I quickly switched to physics and computer science, then later realized
I was a lot more interested in computer science than in physics.  I ended up
with minors in physics and math because of my early forays into physics, and I
think that has been a great preparation for some of the more rigorous material
I have encountered and will encounter in advanced computer science classes and
research.

During my sophomore year, I started working in the Applied Machine Learning lab
with Dr.  Kevin Seppi, with whom I have worked for the past two years.  I moved
through the CS program quickly, leaving me time to take two graduate-level
courses as an undergraduate.  Because I have been moving so quickly, by the
time I get my bachelors degree I will hopefully have two and possibly three
publications.  I was the second author on a conference paper that was accepted
at the Congress on Evolutionary Computation 2009, and I am the first author on
a paper we are submitting to the Genetic and Evolutionary Computation
Conference 2010.  I am also in the process of expanding my conference paper to
a journal paper that we hope to submit to Transactions on Evolutionary
Computation in February.  While the journal paper certainly will be not
published by the time I graduate,  at least it will have been submitted as an
undergraduate paper.

I am applying to the BYU Masters program because I have really enjoyed my time
here as an undergraduate.  I think Dr. Seppi is a great advisor and want to 
keep working with him for my masters degree.  Because of the extra classes I
have taken, I also will have three classes to apply towards a masters degree
that I probably would have difficulty transferring to another school.  As I
already have an advisor and three of the eight classes done, I hope to finish
my masters degree by April 2011.  My current intended area of focus during my
masters degree is natural language processing, and probably unsupervised
morphology in that field.  Because that is my intended topic, I will probably
also be working closely with Dr. Ringger during my stay here.

\end{document}
