\documentclass[onecolumn, 10pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Semester Summary}
\author{}
\date{}

\begin{document}
\maketitle

\subsubsection*{Fall 2004}
\begin{itemize}
  \item Started out double majoring in physics and music.
  \item Quickly switched to double majoring in physics and computer science.
  \item Started off college with 18 credits.
\end{itemize}
\subsubsection*{Winter 2005}
\begin{itemize}
  \item Took an ambitious 21 credits and managed to keep a 4.0.
  \item Received a mission call to Alabama, speaking Spanish.
  \item Member of the Honors Student Advisory Council.
\end{itemize}
\subsubsection*{Summer 2007}
\begin{itemize}
  \item First classes back from the mission, eased back into physics and math.
  \item Started doing physics research for my physics senior thesis.
\end{itemize}
\subsubsection*{Fall 2007}
\begin{itemize}
  \item Tried 18 credits, but I had forgotten too much math, so I dropped a 
	physics class.
  \item Realized I didn't like physics research, and that I wanted to do
	computer science more than physics.
  \item Dropped the physics major to a minor.
  \item Completed all coursework for both a physics and a math minor.
  \item Started working in a computer science research lab with Dr. Kevin Seppi.
\end{itemize}
\subsubsection*{Winter 2008}
\begin{itemize}
  \item Completed several GE courses.
  \item Continued work with Dr. Seppi, doing particle swarm optimization (PSO).
  \item Used PSO to help a physics professor with his crystallography research.
\end{itemize}
\subsubsection*{Spring/Summer 2008}
\begin{itemize}
  \item Presented the PSO/crystallography research at the BYU Spring Research
	Conference.
  \item Took 4 computer science classes and jumped a year ahead in the program.
  \item Began some experiments that led to a publication in a PSO conference.
\end{itemize}
\subsubsection*{Fall 2008}
\begin{itemize}
  \item Did a computer science capstone project with my brother and his friend.
  \item Took my first natural language processing class, starting an interest
	that I plan to pursue in my masters degree.
  \item Met and married my wife---maintained my 4.0 GPA while engaged.
  \item Co-authored a publication accepted at the Congress on Evolutionary
	Computation, a prestigious venue for PSO research.
\end{itemize}
\subsubsection*{Winter 2009}
\begin{itemize}
  \item Took a graduate-level class in natural language processing, which led
	to a submission to the BYU Studies 50th Anniversary Symposium.
  \item Decided on a topic for my honors thesis.
  \item Wrote most of the code for my honors thesis experiments.
\end{itemize}
\subsubsection*{Spring/Summer 2009}
\begin{itemize}
  \item Did a summer internship in California at Google.
\end{itemize}
\subsubsection*{Fall 2009}
\begin{itemize}
  \item Ran all of the experiments I needed for my honors thesis.
  \item Wrote my honors thesis.
  \item Audited an ancient Greek class that I really enjoyed.
  \item Took another graduate-level CS class, in multi-agent systems and game
	theory.
\end{itemize}
\subsubsection*{Winter 2010 (Projected)}
\begin{itemize}
  \item Defended my thesis.
  \item Finished all of my GEs.
  \item My wife had our first child.
  \item I graduated with honors and a 4.0 GPA!
\end{itemize}

\end{document}
