\documentclass[onecolumn, 12pt]{article}
\usepackage{fullpage}
\usepackage{graphicx}

\title{Service Experience}
\author{}
\date{}

\begin{document}
\maketitle

My most memorable service experiences at BYU were in two separate areas.  The
first is volunteer tutoring.  I tutored a number of people over the years, in
physics, math, and computer science.  Tutoring helped my solidify concepts that
I had learned, and it let me share the talents God has given me with those who
needed my help.  One particular example of service in tutoring comes to mind.
My friend James was taking CS 142, the introductory computer science class, and
was struggling quite a bit.  I helped him on a few of the labs earlier in the
semester, but he was having a very difficult time with the last lab.  I found
him in the open labs the evening before the lab was due and told him I could
help him.  The next day he came for help, but it was late and I didn't have
much time.  I helped him for a few minutes then had to go to class.  After
class my wife and I went to the Mormon Tabernacle Choir Christmas Concert with
the Honors program, so I could go help him then.  But I called him and asked if
he needed help, and helped him over the phone for an hour or so, before the
concert.  After the concert I called him again, and he still needed help, so I
talked him through it some more on the way home.  We got back at 11:30 PM, the
lab was due at midnight, and he still hadn't finished.  The bus dropped us off
at the Maeser building, and James was in the Talmage building and was running
out of time.  I ran to the Talmage building and was able to help him finish and
pass off the lab five minutes before it was due.  

I tell that story because the appreciation that James showed to me affected me
a lot.  He was incredibly grateful and said some very nice things, and it
reminded me of the value of giving service.  Not only was James very happy that
he was able to finish his assignment, I was happy because I knew that I had
done something good for someone else.  That experience reinforced for me the
value of giving selfless service and made me want to continue serving.

The second area of service that has been very memorable for me has been in
volunteering at the Missionary Training Center.  My wife served a mission in
Slovenia, and there are very few Slovene speakers in Provo.  The MTC needs
volutneers to come every week so that the missionaries heading to Slovenia have
someone to practice with.  My wife went every week before I met her, and since
we met I have gone with her.  At first I didn't understand what was being said,
but then I started to learn Slovene, in large part so I could help out at the
MTC.  For more than a year I have gone to the MTC almost every week to help the
missionaries practice their teaching and language skills.  This experience was
formative for me because now I speak another language, having learned it
through my service.  It also has been wonderful for me in strengthening my
testimony of and knowledge about the gospel of Jesus Christ.  

\end{document}
